CS182: NBA Scheduler

### 2015-12-08 proposal.tex

	\section{Resources}
	\begin{itemize}
		\item \url{http://www.public.asu.edu/~huanliu/AI04S/project1.htm}
		\item \url{http://www.cc.gatech.edu/~bdilkina/papers/NFLsched1.pdf}
		\item \url{http://www.sloansportsconference.com/wp-content/uploads/2015/02/SSAC15-RP-Finalist-Alleviating-competitive-imbalance-in-NFL-schedules.pdf}
	\end{itemize}

>>>>

	\section{Integer Linear Programming}
	Integer linear programming is a mathematical optimization where all variables are restricted to integers. The objective function and the constraints are linear. In the NBA scheduling problem, the objective function is not defined and the solver has to solve the constraints are formulated in the problem. We solved this problem in 2 steps. The first was to attempt getting a schedule roundwise without considering the dates. Next we included dates and were able to create a player friendly schedule where there no more than 3 games in 5 days.

	\subsection{Methodology for the roundwise schedule}
	We create a variable for every team against every other team, for every round. The first part of the variable is the home team followed by the away team and the round number. We introduce the following constraints to the ILP
	\begin{enumerate}
		\item All teams play 82 games , 41 home and away
		\item All constraints for the division, in conference and out of conference teams
		\item No team can play more than one game in a round
		\item Team can only play either home or away
	\end{enumerate}
	The above ILP was solved using the PuLP as well as the GUROBI solver. The PuLP solver took 40 minutes to converge to solution whereas the GUROBI solver took 6 seconds on an average. The schedule was verified using tests to check if all constraints were satisfied.

	\subsection{Generating the player friendly schedule}
	Using the dates generated by datagen, we formulated the ILP for a player friendly schedule. The variables are created for every date instead of every round.  To reduce the number of variables we generate the variables for every possible home and away dates only. This reduces the number of variables by about 70\%. The time taken to create the constraints and to solve the ILP are greatly reduced. We use the following constraints.
	\begin{enumerate}
		\item All teams play 82 games , 41 home and away
		\item All constraints for the division, in conference and out of conference teams
		\item No team can play more than one game on a single date
		\item Team can only play either home or away
		\item A team can only play 3 or lesser matches in 5 days
		\item A team does not play 2 away games without a day's break
		\item A team does not play an away game right after a home game
	\end{enumerate}
	The above ILP was solved using the PuLP as well as the GUROBI solver. The PuLP solver took 1 hour to converge to solution whereas the GUROBI solver took 29 seconds on an average. In some cases the 4 games in 5 days was unavoidable due to mismatch in dates. But our solution has minimized it to a great extent and can ensure player vitality.
